---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
options(width = 100)
Sys.setlocale("LC_COLLATE", "en_US.UTF-8") # ensure common sorting envir
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
ver <- desc::desc_get_version(".")
ver <- paste0("https://img.shields.io/badge/Version-", ver,
              "-success.svg?style=flat&logo=github")
```


# The `helpr` package

<!-- badges: start -->
![GitHub version](`r ver`)
[![CRAN status](http://www.r-pkg.org/badges/version/helpr)](https://cran.r-project.org/package=helpr)
[![R-CMD-check](https://github.com/stufield/helpr/workflows/R-CMD-check/badge.svg)](https://github.com/stufield/helpr/actions)
[![](https://cranlogs.r-pkg.org/badges/grand-total/helpr)](https://cran.r-project.org/package=helpr)
[![Codecov test coverage](https://codecov.io/gh/stufield/helpr/branch/main/graph/badge.svg)](https://app.codecov.io/gh/stufield/helpr?branch=main)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://choosealicense.com/licenses/mit/)
<!-- badges: end -->



## Overview

The `helpr` package contains numerous helper functions
necessary for generalized machine learning analyses.

It contains and exports numerous utilities and S3 generics
used throughout the my personal analysis suite.
It intentionally favors base R over higher
level *tidyverse* to keep imports to a minimum.

  

----------------


## Installation


```{r install-github, eval = FALSE}
remotes::install_github("stufield/helpr")

# OR

remotes::install_version("stufield/helpr", "0.0.1")
```


----------------


## Usage

To load `helpr` simply make a call to `library()` as usual:

```{r load, eval = FALSE}
library(helpr)
```


## Help summary of the package

```{r help, eval = FALSE}
library(help = helpr)
```


## Notes about the package

* [ ] TODO
* [ ] TODO
* [ ] TODO


