---
title: "Library Mode"
author: "Stu Field"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Library Mode}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(helpr)
library(remotes)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```



# Introduction

* Lack of reproducibility is a major obstacle for any large (or small) organization,
  especially those that deliver products that need an audit trail. Installing
  3rd party packages from external repositories, or internal "between-release"
  development versions of packages, without version or commit tagging
  makes this near to impossible at scale.

* Furthermore, installing a specific package version into a top-level
  environment can make that version the default for all other
  analyses performed in that environment.
  
* `lib_mode()` attempts to solve these issues by creating an
  _specific_ library location (i.e. `lib.loc`) into a
  specified directory at the top of the R library tree.

* Defaults can be easily set via `getOption("helpr_path")`.

* Versions of packages can be installed to this location in a 
  reproducible way (e.g. via the `remotes` package or 
  via `remotes::install_github()` or `remotes::install_version()`),
  which can be scripted and reproduced accurately by subsequent analysts.

* The examples below show:
  - how to use `lib_mode()`
  - how to toggle back-and-forth between modes
  - how to install to the default library mode location

--------------

## Useful Functions

* `lib_mode()`
* `lib_tree()`
* `is_lib_mode()`

------------


## Usage

### Toggle `lib_mode()` ON/OFF

Library Mode usage requires a path specification of the desired library
location via its `path =` argument. This can be _anywhere_ on the file system
that the user has privileges.


```{r lib-mode}
mylib <- "~/tmp-lib"
dir.create(mylib)     # must create a valid directory

lib_tree()

lib_mode(mylib)  # activate; creates lib.loc location

lib_tree()

lib_mode(mylib)  # deactivate

lib_tree()
```


### Install 3rd party packages in `lib_mode()`

For reproducibility, it is critical to install to a `path` using a
specific package version, this can be accomplished via
`remotes::install_version()` if installing a third-party package:

```{r install-in-lib-mode}
packageVersion("spelling")

lib_mode(mylib)

# install old `here` and `spelling` versions
# by default installs to .libPaths()[1L]
install_version("here", version = "0.1", dependencies = FALSE,
                upgrade = FALSE, quiet = TRUE)
install_version("spelling", version = "1.1", dependencies = FALSE,
                upgrade = FALSE, quiet = TRUE)

packageVersion("spelling")

lib_mode(mylib)
```


### Re-activating `lib_mode()`

The next time you activate "lib mode", you will be notified
that you are using a set of installed libraries (`here` and `spelling`):

```{r re-activate}
lib_mode(mylib)
```


-------------


```{r unset, include = FALSE}
lib_mode(mylib)   # turn OFF for below
```

## Using `getOption("helpr_path")`

Another option to use `lib_mode()` without arguments is to set the
`helpr_path` option via `options(helpr_path = "path/to/lib")`. This can be set
early in a script or in a setup chunk of an `Rmarkdown`. It is discouraged
to set this option in an `.Rprofile` because it would not be portable
across users and/or machines.

```{r option-helpr-path}
# uses the 'helpr_path' option set
options(helpr_path = mylib)
lib_mode()
```


```{r cleanup, include = FALSE}
if ( dir.exists(mylib) ) {
  unlink(mylib, recursive = TRUE, force = TRUE)
}
if ( dir.exists("aux-lib") ) {
  unlink("aux-lib", recursive = TRUE, force = TRUE)
}
```
